---
title: "mitochondria_injection_dilution_optimization"
author:
  - name: "Margot Biesemans, Paweł Wiczling"
    affiliations:
      - name: "Department of Biopharmaceutics and Pharmacodynamics, Medical University of Gdańsk, Gen. J. Hallera 107, 80-416 Gdańsk, Poland"
date: "`r format(Sys.Date())`"
format:
  html:
    theme: cosmo
    toc: true
    code-fold: true  
    code-tools: true
    fig-width: 7
    fig-height: 7
---

## Introduction

## Setup

### Knitr settings

```{r knitr-options}
knitr::opts_chunk$set(cache = FALSE, 
                      message = FALSE, 
                      error = FALSE, 
                      warning = FALSE, 
                      comment = NA, 
                      out.width = '95%')
```

### Packages

```{r load-packages}
library(bbr)
library(bbr.bayes)
library(dplyr)
library(ggplot2)
library(patchwork)
library(naniar)
library(knitr)
library(data.table)
library(tidyverse)
library(glue)
library(whisker)
library(here)
library(cmdstanr)
library(gridExtra)
library(posterior)
library(tidybayes)
library(GGally)
library(mrgmisc)
library(scales)

#remotes::install_github("metrumresearchgroup/mrgmisc")
#set_cmdstan_path("C:/Users/Margot/.cmdstan/cmdstan-2.36.0") #change path to your CmdStan location
set.seed(10271998) #setting seed is not required but assures repeatable results
```

### Paths

```{r set-paths}

#data directory
  data_dir = here::here("data", "rawdata")
  #derived data subfolder
    data_derived_dir <- here::here("data", "derived", "mod4") 
    if(!file.exists(data_derived_dir)) dir.create(data_derived_dir, recursive = T)
#model directory
  model_dir <- here::here("model")  
  if(!file.exists(model_dir)) dir.create(model_dir, recursive = T)
#figures and tables directory
  #figures subfolder
    figures_dir <- here::here("deliv", "figures", "mod4")
    if(!file.exists(figures_dir)) dir.create(figures_dir, recursive = T)
  #tables subfolder
    tables_dir <- here::here("deliv", "tables", "mod4")
    if(!file.exists(tables_dir)) dir.create(tables_dir, recursive = T)
```

## Load data

```{r}

data_wide <- read.csv(here(data_dir, 'mito_dilop_stand_pos.csv')) %>% 
  mutate(IS = X237.081234046689.3.68837108833056, .after = time)

data_long_pos <- data_wide %>% 
  tidyr::gather(metabolite_code, intensity, 8:191, factor_key = TRUE) %>%
  rename(chromatogram_name = chrom,
         dilution_factor = dil,
         injection_volume = inj_vol,
         condition = cond,
         replicate = rep,
         chromatogram_number = time,
         internal_standard = IS) %>%
  mutate(metabolite_number =  match(metabolite_code, unique(metabolite_code))) %>%
  filter(intensity != 0) %>%
  mutate(dilution_factor_original = dilution_factor) %>%
  mutate(dilution_factor = if_else(metabolite_number == 184, 
                                   1, dilution_factor)) %>%
  mutate(metabolite_number = 1000 + metabolite_number) %>%
  mutate(log_intensity =  log(intensity),
         log_intensity_factor = log((intensity/(dilution_factor*injection_volume))),
         normalized_intensity = intensity/internal_standard,
         log_normalized_intensity = log(intensity/internal_standard),
         log_normalized_intensity_factor = log(normalized_intensity/(dilution_factor*injection_volume)),
         log_injection_quantity = log(injection_volume*dilution_factor)) %>%
  select(metabolite_number, dilution_factor, dilution_factor_original,
         injection_volume, log_injection_quantity, chromatogram_number, 
         replicate, internal_standard, intensity, log_intensity,
         log_intensity_factor, normalized_intensity, log_normalized_intensity,
         log_normalized_intensity_factor, condition, metabolite_code,
         chromatogram_name) %>%
  mutate(ionization_mode = "positive",
         ionization_code = 1)

data_long_pos$retention_time <- as.double(sub("^([^\\.]*\\.[^\\.]*)\\.(.*)$", "\\2", data_long_pos$metabolite_code))

data_wide <- read.csv(here(data_dir, 'mito_dilop_stand_neg.csv')) %>% 
  mutate(IS = X237.080977847156.3.68490606206505, .after = time)

data_long_neg <- data_wide %>% 
  tidyr::gather(metabolite_code, intensity, 8:909, factor_key = TRUE) %>%
  rename(chromatogram_name = chrom,
         dilution_factor = dil,
         injection_volume = inj_vol,
         condition = cond,
         replicate = rep,
         chromatogram_number = time,
         internal_standard = IS) %>%
  mutate(metabolite_number =  match(metabolite_code, unique(metabolite_code))) %>%
  filter(intensity != 0) %>%
  mutate(dilution_factor_original = dilution_factor) %>%
  mutate(dilution_factor = if_else(metabolite_number == 902, 
                                   1, dilution_factor)) %>%
  mutate(log_intensity = log(intensity),
         log_intensity_factor = log((intensity/(dilution_factor*injection_volume))),
         normalized_intensity = intensity/internal_standard,
         log_normalized_intensity = log(intensity/internal_standard),
         log_normalized_intensity_factor = log(normalized_intensity/(dilution_factor*injection_volume)),
         log_injection_quantity = log(injection_volume*dilution_factor)) %>%
select(metabolite_number, dilution_factor, dilution_factor_original, 
       injection_volume, log_injection_quantity, chromatogram_number, replicate,
       internal_standard, intensity, log_intensity, log_intensity_factor,
       normalized_intensity, log_normalized_intensity,
       log_normalized_intensity_factor, condition, metabolite_code,
       chromatogram_name) %>%
  mutate(ionization_mode = "negative",
         ionization_code = 2)

data_long_neg$retention_time <- as.double(sub("^([^\\.]*\\.[^\\.]*)\\.(.*)$", "\\2", data_long_neg$metabolite_code))

data_long = rbind(data_long_pos, data_long_neg) %>%
  mutate(observation_number = row_number())

rm(data_long_neg, data_long_pos)

write_csv(data_long, file.path(data_derived_dir, "data-posneg.csv"), na = ".")

```

## Running and loading model in cmdstanr

### Run model

```{r, eval =FALSE}
#Execute to read files
#cmdstan_model(here::here(file.path(model_dir, "model", "model.stan")), stanc_options = list("O1"), cpp_options = list(stan_threads = TRUE))
model <- read_model(file.path(model_dir, "model"))
#Execute to build model
model_res <- model %>% submit_model(.overwrite = TRUE)
```

### Generate quantities

```{r, eval=FALSE}
# Execute to get the templates to generate quantities from the model
#model_gq  <- read_model(file.path(model_dir, "model")) %>%
# copy_model_as_stan_gq(.overwrite = TRUE)
# Execcute to read the files
model_gq  <- read_model(file.path(model_dir, "model_gq"))
# Execute to generate quantities
model_gq <- model_gq %>% submit_model(.overwrite = TRUE)
```

### Read the model

```{r, eval = FALSE}
#Execute to read the model and generated quantities results
  model_to_plot  <- read_model(here::here(model_dir, "model")) %>%
    read_fit_model()
  model_to_plot_gq <- read_model(here::here(model_dir, "model_gq")) %>%
    read_fit_model()
```

## Running and loading model in Tryton

```{r paths2}
model <- read_model(file.path(model_dir, "model4"))
model_path = model$absolute_model_path;
model_output <- "model-output"

if(!file.exists(here(model_path,model_output))) {
  dir.create(here(model_path,model_output), recursive = T)}

model_path_gq <- sub("model4$", "model4_gq", model_path)
model_output_gq <- "model_gq-output"
if(!file.exists(here(model_path_gq,model_output_gq))) dir.create(here(model_path_gq,model_output_gq), recursive = T)

decision_path_gq <- sub("model4$", "decision4_gq", model_path)
decision_output_gq <- "decision_gq-output"
if(!file.exists(here(decision_path_gq,decision_output_gq))) dir.create(here(decision_path_gq,decision_output_gq), recursive = T)
```

```{r dump-files, eval = FALSE}
source(paste0(model_path, "/model4-init.R"))
source(paste0(model_path, "/model4-standata.R"))

data <- make_standata()
init <- make_init(.data=data)
write_stan_json(data, here(model_path, "standata.json"), always_decimal = FALSE)
write_stan_json(init(), here(model_path, "init-1.json"), always_decimal = FALSE)
write_stan_json(init(), here(model_path, "init-2.json"), always_decimal = FALSE)
write_stan_json(init(), here(model_path, "init-3.json"), always_decimal = FALSE)
write_stan_json(init(), here(model_path, "init-4.json"), always_decimal = FALSE)
write_stan_json(init(), here(model_path, "init-5.json"), always_decimal = FALSE)
write_stan_json(init(), here(model_path, "init-6.json"), always_decimal = FALSE)
write_stan_json(init(), here(model_path, "init-7.json"), always_decimal = FALSE)
write_stan_json(init(), here(model_path, "init-8.json"), always_decimal = FALSE)
```

```{r read-tryton-files}
model_to_plot <- cmdstanr::as_cmdstan_fit(c(
                                  here(model_path, model_output, 'output_1.csv'),
                                  here(model_path, model_output, 'output_2.csv'),
                                 here(model_path, model_output, 'output_3.csv'),
                                 here(model_path, model_output, 'output_4.csv'),
                                  here(model_path, model_output, 'output_5.csv'),
                                 here(model_path, model_output, 'output_6.csv'),
                                 here(model_path, model_output, 'output_7.csv'),
                                 here(model_path, model_output, 'output_8.csv')
                                ))
```

```{r tryton-gq, eval = FALSE}
#read_model(file.path(model_dir, "model")) %>%  copy_model_as_stan_gq()

source(paste0(model_path_gq, "/model4_gq-fitted-params.R"))
source(paste0(model_path_gq, "/model4_gq-standata.R"))

data <- make_standata()
paths <- c(
    here(model_path, model_output, 'output_1.csv'),
    here(model_path, model_output, 'output_2.csv'),
    here(model_path, model_output, 'output_3.csv'),
    here(model_path, model_output, 'output_4.csv'),
    here(model_path, model_output, 'output_5.csv'),
    here(model_path, model_output, 'output_6.csv'),
    here(model_path, model_output, 'output_7.csv'),
    here(model_path, model_output, 'output_8.csv')
  )

model_gq <- cmdstan_model(here(model_path_gq, "model4_gq.stan"))
model_gq  <- model_gq$generate_quantities(paths,
                                         data = data,
                                         seed = 123,
                                         parallel_chains = 8,
                                         output_basename = "output_gq",
                                         output_dir = here(model_path_gq,model_output_gq))
```

```{r tryton-gq-load-files}
model_to_plot_gq <- cmdstanr::read_cmdstan_csv(c(
                                  here(model_path_gq, model_output_gq, 'output_gq-1.csv'),
                                  here(model_path_gq, model_output_gq, 'output_gq-2.csv'),
                                  here(model_path_gq, model_output_gq, 'output_gq-3.csv'),
                                  here(model_path_gq, model_output_gq, 'output_gq-4.csv'),
                                  here(model_path_gq, model_output_gq, 'output_gq-5.csv'),
                                  here(model_path_gq, model_output_gq, 'output_gq-6.csv'),
                                  here(model_path_gq, model_output_gq, 'output_gq-7.csv'),
                                  here(model_path_gq, model_output_gq, 'output_gq-8.csv')
                                ))
```

### Decision

```{r tryton-gq-dec, eval = FALSE}
#read_model(file.path(model_dir, "model")) %>%  copy_model_as_stan_gq()

source(paste0(decision_path_gq, "/decision4_gq-standata.R"))

data <- make_standata()
paths <- c(
    here(model_path, model_output, 'output_1.csv'),
    here(model_path, model_output, 'output_2.csv'),
    here(model_path, model_output, 'output_3.csv'),
    here(model_path, model_output, 'output_4.csv'),
    here(model_path, model_output, 'output_5.csv'),
    here(model_path, model_output, 'output_6.csv'),
    here(model_path, model_output, 'output_7.csv'),
    here(model_path, model_output, 'output_8.csv')
  )

decision_gq <- cmdstan_model(here(decision_path_gq, "decision4_gq.stan"))
decision_gq  <- decision_gq$generate_quantities(paths,
                                         data = data,
                                         seed = 123,
                                         parallel_chains = 8,
                                         output_basename = "output_gq",
                                         output_dir = here(decision_path_gq,decision_output_gq))
```

```{r tryton-gq-load-files-dec}
decision_to_plot_gq <- cmdstanr::read_cmdstan_csv(c(
                                  here(decision_path_gq, decision_output_gq, 'output_gq-1.csv'),
                                  here(decision_path_gq, decision_output_gq, 'output_gq-2.csv'),
                                  here(decision_path_gq, decision_output_gq, 'output_gq-3.csv'),
                                  here(decision_path_gq, decision_output_gq, 'output_gq-4.csv'),
                                  here(decision_path_gq, decision_output_gq, 'output_gq-5.csv'),
                                  here(decision_path_gq, decision_output_gq, 'output_gq-6.csv'),
                                  here(decision_path_gq, decision_output_gq, 'output_gq-7.csv'),
                                  here(decision_path_gq, decision_output_gq, 'output_gq-8.csv')
                                ))
```

## Model descriptors

### Model parameter summary

```{r}
plot_parameter_tables <- function(model_to_plot, pars){ 
  model_to_plot$summary(c(pars)) %>%
  mutate_if(is.numeric, ~formatC(., 3)) %>% 
  rename(parameter = variable) %>%
  mutate("90% CI" = paste("(", q5, ", ", q95, ")", sep = "")) %>%
  select(parameter, mean, median, sd, mad, "90% CI", ess_bulk, ess_tail, rhat)
}

pars = c("lp__", "mean_metabolite", "mean_dilution", "mean_injection",
         "mean_dilution_injection_interaction", "mean_sigma_dilution",
         "mean_sigma_injection", "mean_sigma_dilution_injection_interaction", 
         "mean_log_sigma", "dmean_metabolite", "dmean_dilution", "dmean_injection",
         "dmean_dilution_injection_interaction", "dmean_sigma_dilution",
         "dmean_sigma_injection", "dmean_sigma_dilution_injection_interaction", 
         "dmean_log_sigma", "sd_metabolite", "sd_dilution", "sd_injection",
         "sd_dilution_injection_interaction", "sd_sigma_dilution",
         "sd_sigma_injection", "sd_sigma_dilution_injection_interaction", 
         "sd_log_sigma", "rho_di[2,1]","rho_ms[2,1]", "sigma_shared")

 #pars = c("z_chrom1")
parameter_table <- plot_parameter_tables(model_to_plot, pars)

parameter_table %>% knitr::kable()

writeLines(knitr::kable(parameter_table %>%
                          select(-mad, -ess_bulk, -ess_tail) %>%
                          filter(str_detect(parameter, "rho", negate = TRUE)), 
                        format = "latex"),  
           here::here(tables_dir, "parameters-table.tex"))

write.csv(parameter_table, file = here::here(tables_dir, "parameter-table.csv"))
```

### Trace plots

```{r}
pars = c("lp__", "mean_metabolite", "mean_dilution", "mean_injection",
         "mean_dilution_injection_interaction", "mean_sigma_dilution",
         "mean_sigma_injection", "mean_sigma_dilution_injection_interaction", 
         "mean_log_sigma", "dmean_metabolite", "dmean_dilution", "dmean_injection",
         "dmean_dilution_injection_interaction", "dmean_sigma_dilution",
         "dmean_sigma_injection", "dmean_sigma_dilution_injection_interaction", 
         "dmean_log_sigma", "sd_metabolite", "sd_dilution", "sd_injection",
         "sd_dilution_injection_interaction", "sd_sigma_dilution",
         "sd_sigma_injection", "sd_sigma_dilution_injection_interaction", 
         "sd_log_sigma", "sigma_shared[1]","sigma_shared[2]")

parameter_labels1 <- c(
  "lp__" = "log_posterior",
  "mean_metabolite" = "mean_metabolite",
  "mean_dilution" = "mean_dilution",
  "mean_injection" = "mean_injection",
  "mean_dilution_injection_interaction" = "mean_interaction",
  "mean_sigma_dilution" = "mean_σ_dilution",
  "mean_sigma_injection" = "mean_σ_injection",
  "mean_sigma_dilution_injection_interaction" = "mean_σ_interaction",
  "mean_log_sigma" = "mean_logσ",
    "dmean_metabolite" = "dmean_metabolite",
  "dmean_dilution" = "dmean_dilution",
  "dmean_injection" = "dmean_injection",
  "dmean_dilution_injection_interaction" = "dmean_interaction",
  "dmean_sigma_dilution" = "dmean_σ_dilution",
  "dmean_sigma_injection" = "dmean_σ_injection",
  "dmean_sigma_dilution_injection_interaction" = "dmean_σ_interaction",
  "dmean_log_sigma" = "dmean_logσ",
  "sd_metabolite" = "SD_metabolite",
  "sd_dilution" = "SD_dilution",
  "sd_injection" = "SD_injection",
  "sd_dilution_injection_interaction" = "SD_interaction",
  "sd_sigma_dilution" = "SD_σ_dilution",
  "sd_sigma_injection" = "SD_σ_injection",
  "sd_sigma_dilution_injection_interaction" = "SD_σ_interaction",
  "sd_log_sigma" = "SD_logσ",
  "sigma_shared[1]" = "sigma_shared1",
    "sigma_shared[2]" = "sigma_shared2"
)
draws_df <- as_draws_df(model_to_plot$draws(pars))

draws_df <- draws_df %>%
  rename_with(~ parameter_labels1[.x], .cols = names(parameter_labels1))

p11 <- bayesplot::mcmc_trace(draws_df) +
  theme(axis.text.x = element_text(size = 6),
        axis.text.y = element_text(size = 6),
        plot.title = element_text(size = 4))

p12 <- bayesplot::mcmc_dens_overlay(draws_df) +
  theme(axis.text.x = element_text(size = 6),
        axis.text.y = element_text(size = 6),
        plot.title = element_text(size = 4))

p11
p12

bayesplot::mcmc_trace(draws_df, pars = c("dmean_dilution"))

ggsave(filename = here::here(figures_dir, "trace-plots.png"), plot = p11, width = 40, height = 15, dpi = 600, units = "cm")
ggsave(filename = here::here(figures_dir, "density-plots.png"), plot = p12, width = 40, height = 15, dpi = 600, units = "cm")
```

### Interval plots

```{r}
pars = c("mean_dilution", "mean_injection", "mean_dilution_injection_interaction", 
         "mean_sigma_dilution", "mean_sigma_injection",   
         "mean_sigma_dilution_injection_interaction", 
         "dmean_dilution", "dmean_injection", "dmean_dilution_injection_interaction", 
         "dmean_sigma_dilution", "dmean_sigma_injection",   
         "dmean_sigma_dilution_injection_interaction",
         "sd_dilution", "sd_injection", "sd_dilution_injection_interaction", 
         "sd_sigma_dilution", "sd_sigma_injection", 
         "sd_sigma_dilution_injection_interaction", 
         "sd_log_sigma")
draws_df <- as_draws_df(model_to_plot$draws(pars))
parameter_labels2 <- intersect(names(draws_df), names(parameter_labels1))
draws_df <- draws_df %>%
  rename_with(~ parameter_labels1[.x], .cols = parameter_labels2)
new_pars <- parameter_labels1[pars]
p21 = bayesplot::mcmc_areas(draws_df, pars = new_pars) 

ggsave(filename = here::here(figures_dir, "parameter-interval-plots.png"), plot = p21, width = 12, height = 10, dpi = 600, units = "cm")

p21
```

### Individual parameters

```{r}

gc()

p301 <- posterior::as_draws_df(model_to_plot$draws("r_metabolite")) %>%
  spread_draws(r_metabolite[metabolite_number]) %>%
  ungroup() %>%
  mutate(ionization_mode = if_else(metabolite_number <= 184, 
                                   "positive", "negative")) %>%
  ggplot(aes(x = r_metabolite, fill = ionization_mode)) +
  geom_histogram(bins = 40, alpha = 0.6, position = "identity") +
  labs(x = expression("metabolite intercept"),
       y = "count") +
  theme(panel.grid = element_blank(),
        axis.ticks.y = element_blank(),      
        axis.text.y = element_text(hjust = 1, size = 10),
        legend.position = "none") 
p302 <- posterior::as_draws_df(model_to_plot$draws("r_metabolite")) %>%
  spread_draws(r_metabolite[metabolite_number]) %>%
  ungroup() %>%
  mutate(ionization_mode = if_else(metabolite_number <= 184, 
                                   "positive", "negative")) %>%
  ggplot(aes(x = ionization_mode, y = r_metabolite, fill = ionization_mode)) +
  geom_violin(alpha = 0.6, trim = FALSE) +
  geom_boxplot(width = 0.1, outlier.size = 0.5) +
  labs(x = NULL, y = expression("metabolite intercept")) +
  theme(panel.grid = element_blank(),
        axis.ticks.y = element_blank(),      
        axis.text.y = element_text(hjust = 1, size = 10)) 
p30 <- p301 + p302 + plot_layout(ncol = 2, widths = c(2, 1))
ggsave(filename = here::here(figures_dir, "metabolite-intercepts.png"),
       plot = p30, width = 14, height = 10, dpi = 2000, units = "cm")
p30

gc()

p31 <- posterior::as_draws_df(model_to_plot$draws("r_metabolite")) %>%
  spread_draws(r_metabolite[metabolite_number]) %>%
  ungroup() %>%
  mutate(ionization_mode = if_else(metabolite_number <= 184, 
                                   "positive", "negative")) %>%
  ggplot(aes(x = r_metabolite, y = reorder(metabolite_number, r_metabolite), 
             color = ionization_mode)) +
  stat_pointinterval(.width = 0.5, size = 2/3) +
  labs(x = expression("metabolite intercept"),
       y = "metabolite number") +
  theme(panel.grid = element_blank(),
        axis.ticks.y = element_blank(),      
        axis.text.y = element_text(hjust = 1, size = 2)) 
ggsave(filename = here::here(figures_dir, "individual-metabolite-intercepts.png"),
       plot = p31, width = 12, height = 10, dpi = 2000, units = "cm")
p31

gc()

p321 <- posterior::as_draws_df(model_to_plot$draws("r_dilution")) %>%
  spread_draws(r_dilution[metabolite_number]) %>%
  ungroup() %>%
  mutate(ionization_mode = if_else(metabolite_number <= 184,
                                     "positive", "negative")) %>%
  ggplot(aes(x = r_dilution, fill = ionization_mode)) +
  geom_histogram(bins = 40, alpha = 0.6, position = "identity") +
  labs(x = expression("dilution slopes"),
       y = "count") +
  theme(panel.grid = element_blank(),
        axis.ticks.y = element_blank(),
        axis.text.y = element_text(hjust = 1, size = 10),
        legend.position = "none") 
p322 <- posterior::as_draws_df(model_to_plot$draws("r_dilution")) %>%
  spread_draws(r_dilution[metabolite_number]) %>%
  ungroup() %>%
  mutate(ionization_mode = if_else(metabolite_number <= 184, 
                                   "positive", "negative")) %>%
  ggplot(aes(x = ionization_mode, y = r_dilution, fill = ionization_mode)) +
  geom_violin(alpha = 0.6, trim = FALSE) +
  geom_boxplot(width = 0.1, outlier.size = 0.5) +
  labs(x = NULL, y = expression("dilution slopes")) +
  theme(panel.grid = element_blank(),
        axis.ticks.y = element_blank(),
        axis.text.y = element_text(hjust = 1, size = 10))
p32 <- p321 + p322 + plot_layout(ncol = 2, widths = c(2, 1))
ggsave(filename = here::here(figures_dir, "dilution-slopes.png"), 
       plot = p32, width = 12, height = 10, dpi = 2000, units = "cm")
p32

gc()

p33 <- posterior::as_draws_df(model_to_plot$draws("r_dilution")) %>%
  spread_draws(r_dilution[metabolite_number]) %>%
  ungroup() %>%
    mutate(ionization_mode = if_else(metabolite_number <= 184,
                                     "positive", "negative")) %>%
  ggplot(aes(x = r_dilution, y = reorder(metabolite_number, r_dilution), 
             color = ionization_mode)) +
  stat_pointinterval(.width = .5, size = 2/3) +
  labs(x = expression("dilution slopes"),
       y = "metabolite number") +
  theme(panel.grid = element_blank(),
        axis.ticks.y = element_blank(),
        axis.text.y = element_text(hjust = 1, size = 2)) 
ggsave(filename = here::here(figures_dir, "individual-dilution-slopes.png"), 
       plot = p33, width = 12, height = 10, dpi = 2000, units = "cm")
p33

gc()

p341 <- posterior::as_draws_df(model_to_plot$draws("r_log_sigma")) %>%
  spread_draws(r_log_sigma[metabolite_number]) %>%
  ungroup() %>%
  mutate(ionization_mode = if_else(metabolite_number <= 184,
                                     "positive", "negative")) %>%
  ggplot(aes(x = r_log_sigma, fill = ionization_mode)) +
  geom_histogram(bins = 40, alpha = 0.6, position = "identity") +
  labs(x = expression("log sigma"),
       y = "count") +
  theme(panel.grid = element_blank(),
        axis.ticks.y = element_blank(),
        axis.text.y = element_text(hjust = 0, size = 10),
        legend.position = "none")
p342 <- posterior::as_draws_df(model_to_plot$draws("r_log_sigma")) %>%
  spread_draws(r_log_sigma[metabolite_number]) %>%
  ungroup() %>%
  mutate(ionization_mode = if_else(metabolite_number <= 184, 
                                   "positive", "negative")) %>%
  ggplot(aes(x = ionization_mode, y = r_log_sigma, fill = ionization_mode)) +
  geom_violin(alpha = 0.6, trim = FALSE) +
  geom_boxplot(width = 0.1, outlier.size = 0.5) +
  labs(x = NULL, y = expression("log sigma")) +
  theme(panel.grid = element_blank(),
        axis.ticks.y = element_blank(),
        axis.text.y = element_text(hjust = 1, size = 10))
p34 <- p341 + p342 + plot_layout(ncol = 2, widths = c(2, 1))
ggsave(filename = here::here(figures_dir, "log-sigma.png"), 
       plot = p34, width = 12, height = 10, dpi = 2000, units = "cm")
p34

gc()

p35 <- posterior::as_draws_df(model_to_plot$draws("r_log_sigma")) %>%
  spread_draws(r_log_sigma[metabolite_number]) %>%
  ungroup() %>%
    mutate(ionization_mode = if_else(metabolite_number <= 184,
                                     "positive", "negative")) %>%
  ggplot(aes(x = r_log_sigma, y = reorder(metabolite_number, r_log_sigma), 
             color = ionization_mode)) +
  stat_pointinterval(.width = .5, size = 2/3) +
  labs(x = expression("log sigma"),
       y = "metabolite number") +
  theme(panel.grid = element_blank(),
        axis.ticks.y = element_blank(),
        axis.text.y = element_text(hjust = 0, size = 2))
ggsave(filename = here::here(figures_dir, "individual-log-sigma.png"), 
       plot = p35, width = 12, height = 10, dpi = 2000, units = "cm")
p35

gc()

p361 <- posterior::as_draws_df(model_to_plot$draws("r_nu")) %>%
  spread_draws(r_nu[metabolite_number]) %>%
  ungroup() %>%
  mutate(ionization_mode = if_else(metabolite_number <= 184,
                                     "positive", "negative")) %>%
  ggplot(aes(x = r_nu, fill = ionization_mode)) +
  geom_histogram(bins = 40, alpha = 0.6, position = "identity") +
  labs(x = expression("nu"),
       y = "count") +
  theme(panel.grid = element_blank(),
        axis.ticks.y = element_blank(),
        axis.text.y = element_text(hjust = 0, size = 10),
        legend.position = "none") 
p362 <- posterior::as_draws_df(model_to_plot$draws("r_nu")) %>%
  spread_draws(r_nu[metabolite_number]) %>%
  ungroup() %>%
  mutate(ionization_mode = if_else(metabolite_number <= 184, 
                                   "positive", "negative")) %>%
  ggplot(aes(x = ionization_mode, y = r_nu, fill = ionization_mode)) +
  geom_violin(alpha = 0.6, trim = FALSE) +
  geom_boxplot(width = 0.1, outlier.size = 0.5) +
  labs(x = NULL, y = expression("nu")) +
  theme(panel.grid = element_blank(),
        axis.ticks.y = element_blank(),
        axis.text.y = element_text(hjust = 1, size = 10))
p36 <- p361 + p362 + plot_layout(ncol = 2, widths = c(2, 1))
ggsave(filename = here::here(figures_dir, "nu.png"), 
       plot = p36, width = 12, height = 10, dpi = 2000, units = "cm")
p36

gc()

p37 <- posterior::as_draws_df(model_to_plot$draws("r_nu")) %>%
  spread_draws(r_nu[metabolite_number]) %>%
    mutate(ionization_mode = if_else(metabolite_number <= 184,
                                     "positive", "negative")) %>%
  ungroup() %>%
  ggplot(aes(x = r_nu, y = reorder(metabolite_number, r_nu), 
             color = ionization_mode)) +
  stat_pointinterval(.width = .5, size = 2/3) +
  labs(x = expression("nu"),
       y = "metabolite number") +
  theme(panel.grid = element_blank(),
        axis.ticks.y = element_blank(),
        axis.text.y = element_text(hjust = 0, size = 2)) 
ggsave(filename = here::here(figures_dir, "individual-nu.png"), 
       plot = p37, width = 12, height = 10, dpi = 2000, units = "cm")
p37

gc()

p381 <- posterior::as_draws_df(model_to_plot$draws("r_injection")) %>%
  spread_draws(r_injection[metabolite_number]) %>%
  ungroup() %>%
  mutate(ionization_mode = if_else(metabolite_number <= 184,
                                     "positive", "negative")) %>%
  ggplot(aes(x = r_injection, fill = ionization_mode)) +
  geom_histogram(bins = 40, alpha = 0.6, position = "identity") +
  labs(x = expression("injection slopes"),
       y = "count") +
  theme(panel.grid = element_blank(),
        axis.ticks.y = element_blank(),
        axis.text.y = element_text(hjust = 0, size = 10),
        legend.position = "none") 
p382 <- posterior::as_draws_df(model_to_plot$draws("r_injection")) %>%
  spread_draws(r_injection[metabolite_number]) %>%
  ungroup() %>%
  mutate(ionization_mode = if_else(metabolite_number <= 184, 
                                   "positive", "negative")) %>%
  ggplot(aes(x = ionization_mode, y = r_injection, fill = ionization_mode)) +
  geom_violin(alpha = 0.6, trim = FALSE) +
  geom_boxplot(width = 0.1, outlier.size = 0.5) +
  labs(x = NULL, y = expression("injection slopes")) +
  theme(panel.grid = element_blank(),
        axis.ticks.y = element_blank(),
        axis.text.y = element_text(hjust = 1, size = 10))
p38 <- p381 + p382 + plot_layout(ncol = 2, widths = c(2, 1))
ggsave(filename = here::here(figures_dir, "injection-slopes.png"), 
       plot = p38, width = 12, height = 10, dpi = 2000, units = "cm")
p38

gc()

p39 <- posterior::as_draws_df(model_to_plot$draws("r_injection")) %>%
  spread_draws(r_injection[metabolite_number]) %>%
  ungroup() %>%
    mutate(ionization_mode = if_else(metabolite_number <= 184,
                                     "positive", "negative")) %>%
  ggplot(aes(x = r_injection, y = reorder(metabolite_number, r_injection), 
             color = ionization_mode)) +
  stat_pointinterval(.width = .5, size = 2/3) +
  labs(x = expression("injection slopes"),
       y = "metabolite number") +
  theme(panel.grid = element_blank(),
        axis.ticks.y = element_blank(),
        axis.text.y = element_text(hjust = 0, size = 2)) 
ggsave(filename = here::here(figures_dir, "individual-injection-slopes.png"), 
       plot = p39, width = 12, height = 10, dpi = 2000, units = "cm")
p39

gc()
```

```{r}
ptest <- posterior::as_draws_df(model_to_plot_gq$generated_quantities,"z_chrom") %>%
  spread_draws(z_chrom[y,posneg]) %>%
  ungroup()%>%
  ggplot(aes(y = z_chrom, x = as.factor(y))) +
  stat_pointinterval(.width = 0.5, size = 2/3) +
  labs(x = expression("var"),
       y = "chrom number") +
  facet_wrap(~posneg)+
  theme(panel.grid = element_blank(),
        axis.ticks.y = element_blank(),      
        axis.text.y = element_text(hjust = 1, size = 4)) 

ptest
```

### Individual histograms

```{r}

gc()

p411 <- posterior::as_draws_df(model_to_plot$draws("r_dilution_injection_interaction")) %>%
  spread_draws(r_dilution_injection_interaction[metabolite_number]) %>%
  ungroup() %>%
  mutate(ionization_mode = if_else(metabolite_number <= 184,
                                     "positive", "negative")) %>%
  ggplot(aes(x = r_dilution_injection_interaction, fill = ionization_mode)) +
  geom_histogram(bins = 40, alpha = 0.6, position = "identity") +
  labs(x = expression("dilution and injection interaction slopes"),
       y = "count") +
  theme(panel.grid = element_blank(),
        axis.ticks.y = element_blank(),
        axis.text.y = element_text(hjust = 0, size = 10),
        legend.position = "none") 
p412 <- posterior::as_draws_df(model_to_plot$draws("r_dilution_injection_interaction")) %>%
  spread_draws(r_dilution_injection_interaction[metabolite_number]) %>%
  ungroup() %>%
  mutate(ionization_mode = if_else(metabolite_number <= 184, 
                                   "positive", "negative")) %>%
  ggplot(aes(x = ionization_mode, y = r_dilution_injection_interaction, fill = ionization_mode)) +
  geom_violin(alpha = 0.6, trim = FALSE) +
  geom_boxplot(width = 0.1, outlier.size = 0.5) +
  labs(x = NULL, y = expression("dilution and injection interaction slopes")) +
  theme(panel.grid = element_blank(),
        axis.ticks.y = element_blank(),
        axis.text.y = element_text(hjust = 1, size = 10))
p41 <- p411 + p412 + plot_layout(ncol = 2, widths = c(2, 1))
ggsave(filename = here::here(figures_dir, "dilution-injection-interaction-slopes.png"), 
       plot = p41, width = 12, height = 10, dpi = 2000, units = "cm")
p41

gc()

p42 <- posterior::as_draws_df(model_to_plot$draws("r_dilution_injection_interaction")) %>%
  spread_draws(r_dilution_injection_interaction[metabolite_number]) %>%
  ungroup() %>%
    mutate(ionization_mode = if_else(metabolite_number <= 184,
                                     "positive", "negative")) %>%
  ggplot(aes(x = r_dilution_injection_interaction, 
             y = reorder(metabolite_number, r_dilution_injection_interaction), 
                  color = ionization_mode)) +
  stat_pointinterval(.width = .5, size = 2/3) +
  labs(x = expression("dilution and injection interaction slopes"),
       y = "metabolite number") +
  theme(panel.grid = element_blank(),
        axis.ticks.y = element_blank(),
        axis.text.y = element_text(hjust = 0, size = 2)) 
ggsave(filename = here::here(figures_dir, "individual-dilution-injection-interaction-slopes.png"), plot = p42, width = 12, height = 10, dpi = 2000, units = "cm")
p42

gc()

p431<-posterior::as_draws_df(model_to_plot$draws("r_metabolite")) %>%
  spread_draws(r_metabolite[metabolite_number]) %>%
  ungroup() %>%
  mutate(ionization_mode = if_else(metabolite_number <= 184,
                                   "positive", "negative")) %>%
  group_by(metabolite_number, ionization_mode) %>%
  summarise(mean_mu = mean(r_metabolite)) %>%
  ggplot(aes(x = mean_mu, fill = ionization_mode)) +
  geom_histogram(alpha = 0.2, position = "identity") +
  labs(x = expression("metabolite intercepts")) 

p432<-posterior::as_draws_df(model_to_plot$draws("r_dilution")) %>%
  spread_draws(r_dilution[metabolite_number]) %>%
  ungroup() %>%
  mutate(ionization_mode = if_else(metabolite_number <= 184,
                                   "positive", "negative")) %>%
  group_by(metabolite_number, ionization_mode) %>%
  summarise(mean_mu = mean(r_dilution)) %>%
  ggplot(aes(x = mean_mu, fill = ionization_mode)) +
  geom_histogram(alpha = 0.2, position = "identity") +
  labs(x = expression("dilution slopes")) 

p433<-posterior::as_draws_df(model_to_plot$draws("r_injection")) %>%
  spread_draws(r_injection[metabolite_number]) %>%
  ungroup() %>%
  mutate(ionization_mode = if_else(metabolite_number <= 184,
                                   "positive", "negative")) %>%
  group_by(metabolite_number, ionization_mode) %>%
  summarise(mean_mu = mean(r_injection)) %>%
  ggplot(aes(x = mean_mu, fill = ionization_mode)) +
  geom_histogram(alpha = 0.2, position = "identity") +
  labs(x = expression("injection slopes")) 

p434<-posterior::as_draws_df(model_to_plot$draws("r_dilution_injection_interaction")) %>%
  spread_draws(r_dilution_injection_interaction[metabolite_number]) %>%
  ungroup() %>%
  mutate(ionization_mode = if_else(metabolite_number <= 184,
                                   "positive", "negative")) %>%
  group_by(metabolite_number, ionization_mode) %>%
  summarise(mean_mu = mean(r_dilution_injection_interaction)) %>%
  ggplot(aes(x = mean_mu, fill = ionization_mode)) +
  geom_histogram(alpha = 0.2, position = "identity") +
  labs(x = expression("dilution and injection interaction slopes")) 

p441<-posterior::as_draws_df(model_to_plot$draws("r_log_sigma")) %>%
  spread_draws(r_log_sigma[metabolite_number]) %>%
  ungroup() %>%
  mutate(ionization_mode = if_else(metabolite_number <= 184,
                                   "positive", "negative")) %>%
  group_by(metabolite_number, ionization_mode) %>%
  summarise(mean_mu = mean(r_log_sigma)) %>%
  ggplot(aes(x = mean_mu, fill = ionization_mode)) +
  geom_histogram(alpha = 0.2, position = "identity") +
  labs(x = expression("logsigmas")) 

p442<-posterior::as_draws_df(model_to_plot$draws("r_nu")) %>%
  spread_draws(r_nu[metabolite_number]) %>%
  ungroup() %>%
  mutate(ionization_mode = if_else(metabolite_number <= 184,
                                   "positive", "negative")) %>%
  group_by(metabolite_number, ionization_mode) %>%
  summarise(mean_mu = mean(r_nu)) %>%
  ggplot(aes(x = mean_mu, fill = ionization_mode)) +
  geom_histogram(alpha = 0.2, position = "identity") +
  labs(x = expression("nu"))

p443<-posterior::as_draws_df(model_to_plot$draws("r_sigma_dilution")) %>%
  spread_draws(r_sigma_dilution[metabolite_number]) %>%
  ungroup() %>%
  mutate(ionization_mode = if_else(metabolite_number <= 184,
                                   "positive", "negative")) %>%
  group_by(metabolite_number, ionization_mode) %>%
  summarise(mean_mu = mean(r_sigma_dilution)) %>%
  ggplot(aes(x = mean_mu, fill = ionization_mode)) +
  geom_histogram(alpha = 0.2, position = "identity") +
  labs(x = expression("dilution sigmas")) 

p444<-posterior::as_draws_df(model_to_plot$draws("r_sigma_injection")) %>%
  spread_draws(r_sigma_injection[metabolite_number]) %>%
  ungroup() %>%
  mutate(ionization_mode = if_else(metabolite_number <= 184,
                                   "positive", "negative")) %>%
  group_by(metabolite_number, ionization_mode) %>%
  summarise(mean_mu = mean(r_sigma_injection)) %>%
  ggplot(aes(x = mean_mu, fill = ionization_mode)) +
  geom_histogram(alpha = 0.2, position = "identity") +
  labs(x = expression("injection sigmas")) 

p445<-posterior::as_draws_df(model_to_plot$draws("r_sigma_dilution_injection_interaction")) %>%
  spread_draws(r_sigma_dilution_injection_interaction[metabolite_number]) %>%
  ungroup() %>%
  mutate(ionization_mode = if_else(metabolite_number <= 184,
                                   "positive", "negative")) %>%
  group_by(metabolite_number, ionization_mode) %>%
  summarise(mean_mu = mean(r_sigma_dilution_injection_interaction)) %>%
  ggplot(aes(x = mean_mu, fill = ionization_mode)) +
  geom_histogram(alpha = 0.2, position = "identity") +
  labs(x = expression("dilution and injection interaction sigmas")) 


p43 <- (p431 + p432)/(p433 + p434) + plot_layout(guides = 'collect')
p44 <- (p441 + p442 + p443)/(p444 + p445) + plot_layout(guides = 'collect')

ggsave(filename = here::here(figures_dir, "individual-histograms-main.png"), 
       plot = p43, width = 16, height = 16, dpi = 600, units = "cm")
ggsave(filename = here::here(figures_dir, "individual-histograms-rest.png"), 
       plot = p44, width = 16, height = 16, dpi = 600, units = "cm")

p43
p44
```

### Parameter correlations

```{r}
data51<-posterior::as_draws_df(model_to_plot$draws("r_metabolite")) %>%
  spread_draws(r_metabolite[metabolite_number]) %>%
  ungroup() %>%
  group_by(metabolite_number) %>%
  summarise(mean_metabolite = mean(r_metabolite))  

data52<-posterior::as_draws_df(model_to_plot$draws("r_dilution")) %>%
  spread_draws(r_dilution[metabolite_number]) %>%
  ungroup() %>%
  group_by(metabolite_number) %>%
  summarise(mean_dilution = mean(r_dilution)) 

data53<-posterior::as_draws_df(model_to_plot$draws("r_injection")) %>%
  spread_draws(r_injection[metabolite_number]) %>%
  ungroup() %>%
  group_by(metabolite_number) %>%
  summarise(mean_injection = mean(r_injection)) 

data54<-posterior::as_draws_df(model_to_plot$draws("r_dilution_injection_interaction")) %>%
  spread_draws(r_dilution_injection_interaction[metabolite_number]) %>%
  ungroup() %>%
  group_by(metabolite_number) %>%
  summarise(mean_interaction = mean(r_dilution_injection_interaction)) 

data55<-posterior::as_draws_df(model_to_plot$draws("r_log_sigma")) %>%
  spread_draws(r_log_sigma[metabolite_number]) %>%
  ungroup() %>%
  group_by(metabolite_number) %>%
  summarise(mean_log_sigma = mean(r_log_sigma)) 

data56<-posterior::as_draws_df(model_to_plot$draws("r_nu")) %>%
  spread_draws(r_nu[metabolite_number]) %>%
  ungroup() %>%
  mutate(ionization_mode = if_else(metabolite_number <= 184,
                                   "positive", "negative")) %>%
  group_by(metabolite_number, ionization_mode) %>%
  summarise(mean_nu = mean(r_nu)) %>%
  select(metabolite_number, mean_nu, ionization_mode)

data57 = cbind(data51, data52, data53, data54, data55, data56)

p51<-ggpairs(data57[, c(2, 4, 6, 8, 10, 12, 13)], columns = c(1:6),
        labeller = "label_parsed",
        aes(color = ionization_mode),
        upper = list(continuous = wrap("cor", size = 2)),
        lower = list(continuous = wrap("points", alpha = 0.3, size = 0.1), 
                 combo = wrap("dot", alpha = 0.4, size = 0.2))) +
theme(axis.text = element_text(size = 5)) +
theme(strip.text.x = element_text(size = 5)) +
theme(strip.text.y = element_text(size = 5))

ggsave(filename = here::here(figures_dir, "individual-pairs-plot.png"), 
       plot = p51, width = 12, height = 16, dpi = 600, units = "cm")

p51


data56 %>% group_by(ionization_mode) %>%
  summarize(mean(as.numeric(mean_nu<=10)))
data56 %>% group_by(ionization_mode) %>%
  summarize(mean(as.numeric(mean_nu<=5)))
```

## Model predictions

```{r}
model_prediction_plots <- function(data_pred_split, data_long_split){
  
p61 <- data_pred_split %>%
  tidyr::expand_grid(dilution_factor = c(0.25, 0.50, 1), injection_volume = c(2, 5,  10)) %>%
  ungroup() %>%
  mutate(logsigma = r_log_sigma + 
           r_sigma_dilution*log(dilution_factor/0.5) + 
           r_sigma_injection*log(injection_volume/5) + 
           r_sigma_dilution_injection_interaction*log(injection_volume/5)*log(dilution_factor/0.5)) %>%
  mutate(ypred1 = r_metabolite +
           r_dilution*log(dilution_factor/0.5) +
           r_injection*log(injection_volume/5) +
           r_dilution_injection_interaction*log(injection_volume/5)*log(dilution_factor/0.5)) %>%
  mutate(ypred2 = r_metabolite +
           r_dilution*log(dilution_factor/0.5) +
           r_injection*log(injection_volume/5) +
           r_dilution_injection_interaction*log(injection_volume/5)*log(dilution_factor/0.5) +
           exp(logsigma)*rt(n(), r_nu)) %>%
  mutate(log_injection_quantity = log(dilution_factor) + log(injection_volume)) %>%
  ggplot(aes(x = log_injection_quantity)) + 
  ggdist::stat_lineribbon(aes(y = ypred2, color = ordered(dilution_factor), 
                              fill = ordered(dilution_factor)), 
                          .width = c(.90), alpha = 0.75) +
  ggdist::stat_lineribbon(aes(y = ypred1, color = ordered(dilution_factor), 
                              fill = ordered(dilution_factor)), 
                          .width = c(.90), alpha = 0.25) +
  geom_jitter(data = data_long_split, aes(x = log_injection_quantity, 
                                        y = log_intensity, 
                                        color = ordered(dilution_factor)), 
                                        alpha = 0.75, width = 0.05, height = 0) +
  facet_wrap(.~metabolite_number_org) +
  labs(x = "Injection quantity (log scale)", y = "Signal intensity (log scale)")

return(p61)
}

xdata <- readr::read_csv(here::here("data", "derived", "data-posneg.csv"))

relabeled_metabolite_numbers <- data.frame(
  metabolite_number = match(xdata$metabolite_number, 
                            unique(xdata$metabolite_number)), 
  metabolite_number_org = xdata$metabolite_number) %>% 
  distinct()

data_pred_split <- posterior::as_draws_df(model_to_plot$draws(c("r_metabolite", "r_dilution", "r_injection", "r_dilution_injection_interaction", "r_log_sigma", "r_sigma_dilution", "r_sigma_injection", "r_sigma_dilution_injection_interaction", "r_nu"))) %>%
  tidybayes::spread_draws(r_metabolite[metabolite_number],
            r_dilution[metabolite_number],
            r_injection[metabolite_number], 
            r_dilution_injection_interaction[metabolite_number],
            r_log_sigma[metabolite_number],
            r_sigma_dilution[metabolite_number],
            r_sigma_injection[metabolite_number],
            r_sigma_dilution_injection_interaction[metabolite_number],
            r_nu[metabolite_number]) %>%
  ungroup() %>%
  left_join(relabeled_metabolite_numbers) %>%
  mutate(PLOTS = ids_per_plot(metabolite_number_org, 9)) %>% # 9 per subplot
  split(.[["PLOTS"]])

data_long_x <- data_long
data_long_x$metabolite_number_org <- data_long_x$metabolite_number

data_long_split <- data_long_x %>%
  mutate(PLOTS = ids_per_plot(metabolite_number_org, 9)) %>% # 9 per subplot
  split(.[["PLOTS"]])

plots_by_id <- purrr::map2(data_pred_split, 
                          data_long_split, 
                          function(x, y) model_prediction_plots(x, y))
rm(data_long_x)

# save to single pdf file
ggsave(
    filename = "fits.pdf",
    plot = marrangeGrob(plots_by_id, nrow = 1, ncol = 1, top = NULL),
    path = figures_dir)
  
# few examples  
plots_by_id[20]
plots_by_id[23]
plots_by_id[121]
```

## Residuals

### Goodness of fit

```{r}
temp <- posterior::as_draws_df(model_to_plot_gq$generated_quantities)

Y_predicted <- apply(temp[, which(colnames(temp) %in% grep("^Ypred", names(temp), value = TRUE))], MARGIN = 2, FUN = quantile, probs = c(.05, .5, .95))

weighted_residuals <- apply(temp[, which(colnames(temp) %in% grep("^wres", names(temp), value = TRUE))], MARGIN = 2, FUN = quantile, probs = c(.05, .5, .95))

Y_predicted <- as.data.frame(t(Y_predicted))
weighted_residuals <- as.data.frame(t(weighted_residuals))

data_long$median_Y_predicted = Y_predicted$`50%`
data_long$median_weighted_residuals = weighted_residuals$`50%`

p71 <- ggplot(data_long, aes(x = log_intensity,
                             y = median_Y_predicted, 
                             color = as.factor(metabolite_number))) + 
  geom_point(size = 0.5, alpha = 0.25) +
   xlab("Observed Y") +
   ylab("Predicted Y") +
   theme(legend.position = "none") +
   geom_abline(intercept = 0, slope = 1, color = "grey", linetype = "dotted")

p72 <- ggplot(data_long, aes(x = chromatogram_number,
                             y = median_weighted_residuals, 
                             color = as.factor(metabolite_number))) + 
   geom_point(size = 0.5, alpha = 0.25) +
   xlab("Measurement order") +
   ylab("Weighted Residuals") +
   theme(legend.position = "none") +
   geom_abline(intercept = 0, slope = 0, color = "grey", linetype = "dotted")

p73 <- ggplot(data_long, aes(x = median_Y_predicted,
                             y = median_weighted_residuals, 
                             color = as.factor(metabolite_number))) + 
   geom_point(size = 0.5, alpha = 0.25) +
   xlab("Predictions") +
   ylab("Weighted Residuals") +
   theme(legend.position = "none") +
   geom_abline(intercept = 0, slope = 0, color = "grey", linetype = "dotted")

p74 <- ggplot(data_long, aes(x = log_injection_quantity,
                             y = median_weighted_residuals, 
                             color = as.factor(metabolite_number))) + 
   geom_point(size = 0.5, alpha = 0.25) +
   xlab("log(injection*dilution") +
   ylab("Weighted Residuals") +
   theme(legend.position = "none") +
   geom_abline(intercept = 0, slope = 0, color = "grey", linetype = "dotted")

pall71 <- (p71 + p72)/(p73 + p74)

ggsave(filename = here::here(figures_dir, "goodness_of_fit.png"), 
       plot = pall71, width = 12, height = 16, dpi = 600, units = "cm")

pall71
```

### Residuals by ID

```{r}
plot_residuals_by_id1 <- function(data_residuals_split){
  
plot <- data_residuals_split %>%
  ggplot(aes(x = log_injection_quantity, 
             y = median_weighted_residuals, 
             shape = as.factor(replicate),  
             color = as.factor(injection_volume))) +
  geom_jitter(height = 0, width = 0.025) +
  labs(x = expression("injection quantity(injection volume * dilution factor on log scale)"),
       y = "weighted residuals",
       color = "injection_volume",
       shape = "replicate") +
  theme_bw() + 
  facet_wrap(.~metabolite_number)

return(plot)
}

data_residuals_split <- data_long %>%
  mutate(PLOTS = ids_per_plot(metabolite_number, 9)) %>% # 9 per subplot
  split(.[["PLOTS"]])

residuals_by_id <- purrr::map(data_residuals_split, 
                          function(x) plot_residuals_by_id1(x))

# save to single pdf file
ggsave(
    filename = "residuals1.pdf",
    plot = marrangeGrob(residuals_by_id, nrow = 1, ncol = 1, top = NULL),
    path = figures_dir)
  
# few examples  
residuals_by_id[1]
residuals_by_id[3]

plot_residuals_by_id2 <- function(data_residuals_split){
  
plot <- data_residuals_split %>%
  ggplot(aes(x = chromatogram_number, y = median_weighted_residuals, shape = as.factor(replicate),  color = as.factor(injection_volume))) +
  geom_jitter(height = 0, width = 0.025) +
  labs(x = expression("order of acquisition"),
       y = "weighted_residuals",
       color = "injection volume",
       shape = "replicate") +
  theme_bw() + 
  facet_wrap(.~metabolite_number)

return(plot)
}

residuals_by_id2 <- purrr::map(data_residuals_split, 
                          function(x) plot_residuals_by_id2(x))

# save to single pdf file
ggsave(
    filename = "residuals2.pdf",
    plot = marrangeGrob(residuals_by_id2, nrow = 1, ncol = 1, top = NULL),
    path = figures_dir)
  
residuals_by_id2[21]
residuals_by_id2[121]
```

```{r}
# Median weighted residuals by dilution factor
  p81 <- data_long %>%
    group_by(chromatogram_number, 
             dilution_factor, 
             replicate, 
             injection_volume, 
             ionization_mode) %>%
    summarise(median_median_weighted_residuals = 
                median(median_weighted_residuals), .groups = "drop") %>%
    ggplot(aes(x = dilution_factor, 
               y = median_median_weighted_residuals, 
               shape = as.factor(replicate),
               color = as.factor(injection_volume))) +
    geom_jitter(height = 0, 
                width = 0.025) +
    labs(x = "Dilution factor",
         y = "Median of median weighted residuals",
         color = "Injection volume",
         shape = "Replicate") +
    theme_bw() +
    facet_wrap(.~ionization_mode)
  
  p81

#Median residuals by chromatogram number
  p82 <- data_long %>%
    group_by(chromatogram_number, 
             dilution_factor, 
             replicate, 
             injection_volume, 
             ionization_mode) %>%
    summarise(median_median_weighted_residuals = 
                median(median_weighted_residuals), .groups = "drop") %>%
    ggplot() +
    geom_point(aes(x = chromatogram_number, 
                   y = median_median_weighted_residuals, 
                   shape = as.factor(injection_volume), 
                   color = as.factor(dilution_factor))) +
    geom_line(aes(x = chromatogram_number, 
                  y = median_median_weighted_residuals), 
                  color = "gray") +
    labs(x = "Chromatogram number",
         y = "Median of median weighted residuals",
         color = "Dilution factor",
         shape = "Injection volume") +
    theme_bw() +
    facet_wrap(.~ionization_mode)
  
  p82

#Median residuals by internal standard
  p83 <- data_long %>%
    group_by(chromatogram_number, 
             dilution_factor, 
             replicate, 
             injection_volume, 
             internal_standard, 
             ionization_mode) %>%
    summarize(median_median_weighted_residuals = 
                median(median_weighted_residuals), .groups = "drop") %>%
    mutate(log_internal_standard = 
             log(internal_standard)) %>%
    ggplot() +
    geom_point(aes(x = log_internal_standard, 
                   y = median_median_weighted_residuals, 
                   shape = as.factor(injection_volume), 
                   color = as.factor(dilution_factor))) +
    geom_line(aes(x = log_internal_standard, 
                  y = median_median_weighted_residuals),
                  color = "gray") +
    labs(x = "Log of internal standard intensity",
         y = "Median of median weighted residuals of internal standard",
         color = "dilution factor",
         shape = "injection volume") +
    theme_bw() +
    facet_wrap(.~ionization_mode)
  
  p83

#Save the plots
  ggsave(filename = here::here(figures_dir,
                               "median-weighted-residuals-dilution.png"), 
         plot = p81, width = 12, height = 16, dpi = 600, units = "cm")
  ggsave(filename = here::here(figures_dir, 
                               "median-weighted-residuals-order.png"), 
         plot = p82, width = 12, height = 16, dpi = 600, units = "cm")
  ggsave(filename = here::here(figures_dir,
                               "median-weighted-residuals-internal-standard.png"), 
         plot = p83, width = 12, height = 16, dpi = 600, units = "cm")
```

```{r}
plot <- posterior::as_draws_df(model_to_plot$draws(c("r_metabolite", "r_dilution", "r_injection", "r_dilution_injection_interaction", "r_log_sigma", "r_nu"))) %>%
  tidybayes::spread_draws(r_metabolite[metabolite_number], 
                          r_dilution[metabolite_number], 
                          r_injection[metabolite_number],
                          r_dilution_injection_interaction[metabolite_number],
                          r_log_sigma[metabolite_number], 
                          r_nu[metabolite_number]) %>%
  ungroup() %>%
  tidyr::expand_grid(dilution_factor = seq(0.25, 1, length.out = 6),
                     injection_volume = c(2, 5, 10)) %>%
  ungroup() %>%
  mutate(ionization_mode = if_else(metabolite_number <= 184, 
                                   "positive", "negative")) %>%
  mutate(Y_predicted = r_metabolite + 
           r_dilution*log(dilution_factor/0.5) +
           r_injection*log(injection_volume/5) +
           r_dilution_injection_interaction*log(injection_volume/5)*log(dilution_factor/0.5)) %>%
  group_by(metabolite_number, 
           dilution_factor, 
           injection_volume, 
           ionization_mode) %>%
  summarise(mean_Y_predicted = mean(Y_predicted)) %>%
  ggplot(aes(x = dilution_factor, y = mean_Y_predicted, color = as.factor(metabolite_number))) +
  geom_line() +
  facet_grid(injection_volume~ionization_mode) +
  labs(x = "Dilution factor", 
       y = "Predicted intensity (log scale)") +
  theme(legend.position = "none")

plot

ggsave(filename = here::here(figures_dir, "simulations.png"), plot = plot, width = 12, height = 16, dpi = 600, units = "cm")

```

# Application

```{r}
#Generate simulations matrix
  gc() 
  simulation_data <- posterior::as_draws_df(model_to_plot$draws(c("r_metabolite", 
                                            "r_dilution",  
                                            "r_injection",  
                                            "r_dilution_injection_interaction",  
                                            "r_log_sigma",  
                                            "r_nu"))) %>% 
    tidybayes::spread_draws(r_metabolite[metabolite_number],  
                            r_dilution[metabolite_number],  
                            r_injection[metabolite_number], 
                            r_dilution_injection_interaction[metabolite_number], 
                            r_log_sigma[metabolite_number],  
                            r_nu[metabolite_number])  
  gc()
    simulation_data <- simulation_data %>%
    ungroup() %>%
    tidyr::expand_grid(dilution_factor = seq(0.25, 1, length.out = 10),
                       injection_volume = seq(2.5, 10, length.out = 10)) %>%
    mutate(ionization_mode = if_else(metabolite_number <= 184,
                                     "positive", "negative")) %>%
    mutate(log_dilution = log(dilution_factor/0.5)) %>%
    mutate(log_injection = log(injection_volume/5)) %>%
    mutate(slope_dilution = r_dilution +
             r_dilution_injection_interaction*log_injection) %>%
    mutate(slope_injection =  r_injection +
             r_dilution_injection_interaction*log_dilution)
  gc()
#  write.csv(simulation_data, file = here::here(data_derived_dir, "simulation_data.csv"))

#Evaluation of slopes
  p91 <- simulation_data %>%
    group_by(metabolite_number, log_dilution, log_injection, ionization_mode) %>%
    summarise(mean_slope_dilution = mean(slope_dilution),
              mean_slope_injection = mean(slope_injection)) %>%
    group_by(log_dilution, log_injection, ionization_mode) %>%
    summarise(mean_of_means_slope_dilution = mean(mean_slope_dilution),
              mean_of_means_slope_injection = mean(mean_slope_injection),
              magnitude = sqrt(mean_of_means_slope_dilution^2 +
                                 mean_of_means_slope_injection^2)) %>%
    ggplot(aes(x = log_dilution, y = log_injection)) +
    geom_segment(aes(xend = log_dilution + 0.1*mean_of_means_slope_dilution,
                     yend = log_injection + 0.1*mean_of_means_slope_injection,
                     color = magnitude),
    arrow = arrow(length = unit(0.15, "cm")), size = 0.5) +
    scale_color_viridis_c() +
    labs(x = "Dilution (log scale)", y = "Injection Volume (log scale)") +
    facet_grid(.~ionization_mode)
  p91
  ggsave(filename = here::here(figures_dir, "simulations-slope.png"),  

         plot = p91, width = 12, height = 16, dpi = 600, units = "cm")
  gc() 
```

# Decision making

```{r}
dilution_factor_sim = seq(0.25, 1, length.out = 10)
injection_volume_sim = seq(2.5, 10, length.out = 10)
posneg = c("positive", "negative")

performance_simulation <- posterior::as_draws_df(
  decision_to_plot_gq$generated_quantities) %>%
  tidybayes::spread_draws(
    SVR[i,j,c],pSVR[i,j,c]) %>%
  ungroup() 

design_summary <- performance_simulation %>%
  group_by(i,j,c) %>%
  summarise(
    mean_metric = mean(SVR),
    proportion_metabolites = mean(pSVR),
    .groups = "drop")

design_summary$dilution_factor_sim = dilution_factor_sim[design_summary$i]
design_summary$injection_volume_sim = injection_volume_sim[design_summary$j]
design_summary$ionization_mode = posneg[design_summary$c]

p_mean <- ggplot(
  design_summary,
  aes(x = dilution_factor_sim, y = injection_volume_sim, z = mean_metric)) +
  geom_contour_filled(bins = 10) +
  facet_wrap(~ ionization_mode) +
  #scale_fill_viridis_c(name = "Mean\nmetric") +
  labs(
    x = "Dilution factor",
    y = "Injection volume",
    fill = "SVR",
    title = "Sensitivity-to-variability ratio")

p_prop <- ggplot(
  design_summary,
  aes(x = dilution_factor_sim, 
      y = injection_volume_sim, 
      z = proportion_metabolites)) +
  geom_contour_filled(bins = 10) +
  facet_wrap(~ ionization_mode) +
  labs(
    x = "Dilution factor",
    y = "Injection volume",
    fill = "pSVR",
    title = "Proportion of Features Quantifiable at Two-Fold Difference")

p_perf_2x2 <- p_mean / p_prop + plot_layout(guides = "collect")

p_perf_2x2

ggsave(
  filename = here::here(figures_dir, "performance-metric-grid.png"),
  plot     = p_perf_2x2,
  width    = 18,
  height   = 18,
  dpi      = 600,
  units    = "cm")
```

```{r}
target_dilutions  <- c(0.25, 0.5, 1.0)
target_injections <- c(2.5, 5, 10)

levels_new = c(
  "Dil = 0.25, Inj = 10",
  "Dil = 0.5, Inj = 10",
"Dil = 1, Inj = 10",
"Dil = 0.25, Inj = 5",
"Dil = 0.5, Inj = 5",
"Dil = 1, Inj = 5",  
"Dil = 0.25, Inj = 2.5",
"Dil = 0.5, Inj = 2.5",
"Dil = 1, Inj = 2.5")

#

performance_subset <- posterior::as_draws_df(
  decision_to_plot_gq$generated_quantities) %>%
  tidybayes::spread_draws(
    svrm[i,j,m]) %>%
  ungroup() 


performance_subset <- performance_subset %>%
  mutate(ionization_mode = if_else(m <= 184, 
                                   "positive", "negative"))

#
performance_pos <- performance_subset %>%
  dplyr::filter(ionization_mode == "positive")

performance_pos_hist = performance_pos %>%
  group_by(i,j,m) %>%
  summarise(mean_performance_metric = mean(svrm))

performance_pos_hist$dilution_factor = target_dilutions[performance_pos_hist$i]
performance_pos_hist$injection_volume = target_injections[performance_pos_hist$j]

 performance_pos_hist <- performance_pos_hist %>%
  mutate(
    design_label = glue("Dil = {dilution_factor}, Inj = {injection_volume}"),
    design_label = factor(design_label, levels = levels_new))
#
counts_pos <- performance_pos %>%
  group_by(i,j,.draw) %>%
  summarise(
    n_total = n(),
    n_above = sum(svrm > 2/log(2)),
    prop_above = n_above/n_total,
    .groups = "drop") %>% 
  group_by(i,j) %>%
   summarise(
    n_total = mean(n_total),
    n_above = mean(n_above),
    prop_above = mean(prop_above),
    .groups = "drop") %>%
  mutate(
    label = paste0(round(n_above,0),
      " (", percent(prop_above, accuracy = 1), ")"))
#

counts_pos$dilution_factor = target_dilutions[counts_pos$i]
counts_pos$injection_volume = target_injections[counts_pos$j]

 counts_pos <- counts_pos %>%
  mutate(
    design_label = glue("Dil = {dilution_factor}, Inj = {injection_volume}"),
    design_label = factor(design_label, levels = levels_new))
  
performance_histogram_pos <- ggplot(
  performance_pos_hist,
  aes(x = mean_performance_metric)) +
  geom_histogram(bins = 30, alpha = 0.7, color = "black") +
  geom_vline(xintercept = 2/log(2), linetype = "dashed", color = "red") +
  facet_wrap(~ design_label, ncol = 3, scales = "fixed") +
  geom_text(
    data = counts_pos,
    aes(x = -Inf, y = Inf, label = label),
    hjust = -0.05, vjust = 1.1,
    inherit.aes = FALSE,
    size = 3) +
  labs(
    x = "Sensitivity-to-variability ratio",
    y = "Number of metabolites",
    title = "Positive ionization mode:\nperformance metric across metabolites")
#
performance_histogram_pos
#
ggsave(
  filename = here::here(figures_dir, "performance-histograms-positive.png"),
  plot     = performance_histogram_pos,
  width    = 18,
  height   = 12,
  dpi      = 600,
  units    = "cm"
)
#
#

performance_neg <- performance_subset %>%
  filter(ionization_mode == "negative")

performance_neg_hist = performance_neg %>%
  group_by(i,j,m) %>%
  summarise(mean_performance_metric = mean(svrm))

performance_neg_hist$dilution_factor = target_dilutions[performance_neg_hist$i]
performance_neg_hist$injection_volume = target_injections[performance_neg_hist$j]

 performance_neg_hist <- performance_neg_hist  %>%
  mutate(
    design_label = glue("Dil = {dilution_factor}, Inj = {injection_volume}"),
    design_label = factor(design_label, levels = levels_new))
 
#
counts_neg <- performance_neg %>%
  group_by(i,j,.draw) %>%
  summarise(
    n_total = n(),
    n_above = sum(svrm > 2/log(2)),
    prop_above = n_above/n_total,
    .groups = "drop") %>% 
  group_by(i,j) %>%
   summarise(
    n_total = mean(n_total),
    n_above = mean(n_above),
    prop_above = mean(prop_above),
    .groups = "drop") %>%
  mutate(
    label = paste0(round(n_above,0),
      " (", percent(prop_above, accuracy = 1), ")"))
#

counts_neg$dilution_factor = target_dilutions[counts_neg$i]
counts_neg$injection_volume = target_injections[counts_neg$j]

 counts_neg <- counts_neg  %>%
  mutate(
    design_label = glue("Dil = {dilution_factor}, Inj = {injection_volume}"),
    design_label = factor(design_label, levels = levels_new))
 
performance_histogram_neg <- ggplot(
  performance_neg_hist,
  aes(x = mean_performance_metric)) +
  geom_histogram(bins = 30, alpha = 0.7, color = "black") +
  geom_vline(xintercept = 2/log(2), linetype = "dashed", color = "red") +
  facet_wrap(~ design_label, ncol = 3, scales = "fixed") +
  geom_text(
    data = counts_neg,
    aes(x = -Inf, y = Inf, label = label),
    hjust = -0.05, vjust = 1.1,
    inherit.aes = FALSE,
    size = 3) +
  labs(
    x = "Sensitivity-to-variability ratio",
    y = "Number of metabolites",
    title = "Negative ionization mode:\nperformance metric across metabolites")
#
performance_histogram_neg
#
ggsave(
  filename = here::here(figures_dir, "performance-histograms-negative.png"),
  plot     = performance_histogram_neg,
  width    = 18,
  height   = 12,
  dpi      = 600,
  units    = "cm"
)
```
